### 2016-05-09 Manuscript/SW_early_life_telomeres.Rmd
More home edits

#####Early-life telomere shortening
Longitudinal data showed that early-life RTL was not related to adult RTL (R^2^ = `r round2(summary(lm(RTL1~RTL+Agemonths,data=Loss))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=Loss))$coefficients[2,3])`; P = `r round2(summary(lm(RTL1~RTL+Agemonths,data = Loss))$coefficients[2,4])`; Fig. 4A). Running this analysis separetly for juvenile age classes revealed that this relationship was not significant for nestlings (R^2^ = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 1)))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 1)))$coefficients[2,3])`; P = `r round2(summary(lm(RTL1~RTL+Agemonths,data = subset(Loss,Agemonths == 1)))$coefficients[2,4])`), fledglings (R^2^ = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 6)))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 6)))$coefficients[2,3])`; P = `r round2(summary(lm(RTL1~RTL+Agemonths,data = subset(Loss,Agemonths == 6)))$coefficients[2,4])`) or subadults (R^2^ = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 10)))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 10)))$coefficients[2,3])`; P = `r round2(summary(lm(RTL1~RTL+Agemonths,data = subset(Loss,Agemonths == 10)))$coefficients[2,4])`). $\Delta$RTL was significantly lower than zero in nestlings (t = `r round2(t.test(subset(Loss,Agemonths == 1)$Loss)$statistic)`; P `r round3(t.test(subset(Loss,Agemonths == 1)$Loss)$p.value)`), but not in fledglings (t = `r round2(t.test(subset(Loss,Agemonths == 6)$Loss)$statistic)`; P = `r round2(t.test(subset(Loss,Agemonths == 6)$Loss)$p.value)`) or subadults (t = `r round2(t.test(subset(Loss,Agemonths == 10)$Loss)$statistic)`; P = `r round2(t.test(subset(Loss,Agemonths == 10)$Loss)$p.value)`). In line with this, $\Delta$RTL increased with age in early life (R^2^ = `r round2(summary(lm(Loss~Agemonths,data=Loss))$r.squared)`; t = `r round2(summary(lm(Loss~Agemonths,data=Loss))$coefficients[2,3])`; P `r round3(summary(lm(Loss~Agemonths,data = Loss))$coefficients[2,4])`; Fig. 4B), suggesting that telomere shortening was highest shortly after the nestling phase.

>>>>

#####Early-life telomere shortening
Longitudinal data showed that early-life RTL was not related to adult RTL (R^2^ = `r round2(summary(lm(RTL1~RTL+Agemonths,data=Loss))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=Loss))$coefficients[2,3])`; P = `r round2(summary(lm(RTL1~RTL+Agemonths,data = Loss))$coefficients[2,4])`; Fig. 4A). Running this analysis separetly for juvenile age classes revealed that this relationship was not significant for nestlings (R^2^ = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 1)))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 1)))$coefficients[2,3])`; P = `r round2(summary(lm(RTL1~RTL+Agemonths,data = subset(Loss,Agemonths == 1)))$coefficients[2,4])`), fledglings (R^2^ = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 6)))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 6)))$coefficients[2,3])`; P = `r round2(summary(lm(RTL1~RTL+Agemonths,data = subset(Loss,Agemonths == 6)))$coefficients[2,4])`) or subadults (R^2^ = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 10)))$r.squared)`; t = `r round2(summary(lm(RTL1~RTL+Agemonths,data=subset(Loss,Agemonths == 10)))$coefficients[2,3])`; P = `r round2(summary(lm(RTL1~RTL+Agemonths,data = subset(Loss,Agemonths == 10)))$coefficients[2,4])`). $\Delta$RTL was significantly lower than zero in nestlings (t = `r round2(t.test(subset(Loss,Agemonths == 1)$Loss)$statistic)`; P `r round3(t.test(subset(Loss,Agemonths == 1)$Loss)$p.value)`), but not in fledglings (t = `r round2(t.test(subset(Loss,Agemonths == 6)$Loss)$statistic)`; P = `r round2(t.test(subset(Loss,Agemonths == 6)$Loss)$p.value)`) or subadults (t = `r round2(t.test(subset(Loss,Agemonths == 10)$Loss)$statistic)`; P = `r round2(t.test(subset(Loss,Agemonths == 10)$Loss)$p.value)`). In line with this, $\Delta$RTL increased with age in early life (R^2^ = `r round2(summary(lm(Loss~Agemonths,data=Loss))$r.squared)`; t = `r round2(summary(lm(Loss~Agemonths,data=Loss))$coefficients[2,3])`; P `r round3(summary(lm(Loss~Agemonths,data = Loss))$coefficients[2,4])`), suggesting that telomere shortening was highest shortly after the nestling phase, after which both telomere shortening and telomere lenghtening were observed (Fig. 4B).

To test whether the apparent telomere lengthening in our dataset could be explained by measurement error, we used the entire Seychelles warbler telomere dataset to compare variance in telomere length among repeat measurements of samples (N = XX samples) to the variance observed among different samples of the same individual (see supplementary material XX for details). We found significantly higher variance among individuals compared to among samples (Levene's test XXX). Importantly, splitting the longitudinal data revealed that not only did we observe significantly more shortening within individuals compared to within samples (XXX), but also significantly more lengthening (XXX).
